---
title: "GEO_tabeller"
author: "Vegard"
date: "2023-10-13"
output: html_document
---

Script to compare and update the GEO-tables in ORGDATA and KHELSA


```{r setup, include=FALSE}
library(orgdata)
library(RODBC)
library(data.table)

# Sett encoding for R >= 4.2

if(as.numeric(R.version$major) >= 4 & as.numeric(R.version$minor) >= 2){
    Sys.setlocale("LC_ALL", "nb-NO.UTF-8")
}
```

# Update GEO-tables for ORGDATA

The following codes are used to generate updated versions of the geo-tables in `Orgdata` (geo-koders.accdb)

- Saves tables in environment for inspection, check this before writing anything to the database
- Change `year` to latest
- Change `write` to TRUE to update `recoding tables` in database
- Change `append` to TRUE to update `tblGeo` in database

```{r}
year <- 2024
write <- FALSE # only applies to recoding tables below, change to TRUE to update in database
append <- FALSE # only applies to tblGeo below, change to TRUE to update in database

# Recoding tables
assign(paste0("kommune", year), orgdata::geo_recode("k", from = 1990, to = year, write = write, append = FALSE))
assign(paste0("fylke", year), orgdata::geo_recode("f", from = 1990, to = year, write = write, append = FALSE))
assign(paste0("grunnkrets", year), orgdata::geo_recode("g", from = 1990, to = year, write = write, append = FALSE))

# tblGeo
tblGeo <- geo_map(year = aargang, write = FALSE, append = append)
```

# Update GEO-tables for KHELSA

The following codes are used to generate updated versions of the tables `GeoKoder` and `KnrHarm` in `KHELSA` (KHELSA.mdb)

## Connection to KHELSA and geo-koder database

- Read relevant tables
- convert all geo-columns to character and add leading 0

```{r}
year <- 2024
KHELSA <- RODBC::odbcConnectAccess2007("F:/Forskningsprosjekter/PDB 2455 - Helseprofiler og til_/PRODUKSJON/STYRING/KHELSA.mdb")
GEOtables <- RODBC::odbcConnectAccess2007("F:/Forskningsprosjekter/PDB 2455 - Helseprofiler og til_/PRODUKSJON/STYRING/raw-khelse/geo-koder.accdb")


# Function to convert selected columns to character and add leading 0
addleading0 <- function(data, cols){
    
    data[, (cols) := lapply(.SD, as.character), .SDcols = cols]
    
    for(i in 1:length(cols)){
        data[nchar(get(cols[i])) %in% c(1,3,5), (cols[i]) := paste0("0", get(cols[i]))]
    }
    
    data[]
}

# Read relevant tables and apply `addleading0()`

## GeoKoder
GeoKoder <- setDT(sqlQuery(KHELSA, "SELECT * FROM GeoKoder"))
addleading0(GeoKoder, "GEO")
## KnrHarm
KnrHarm <- setDT(sqlQuery(KHELSA, "SELECT * FROM KnrHarm"))
addleading0(KnrHarm, c("GEO", "GEO_omk"))

## kommune and fylke
kommune <- setDT(sqlQuery(GEOtables, paste0("SELECT oldCode, currentCode, changeOccurred FROM kommune", year)))
addleading0(kommune, c("oldCode", "currentCode"))

fylke <- setDT(sqlQuery(GEOtables, paste0("SELECT oldCode, currentCode, changeOccurred FROM fylke", year)))
addleading0(fylke, c("oldCode", "currentCode"))

## tblGeo
tblGeo <- setDT(sqlQuery(GEOtables, paste0("SELECT * FROM tblGeo WHERE validTo = '", year, "' AND level <> 'grunnkrets'")))
addleading0(tblGeo, c("code", "grunnkrets", "kommune", "fylke", "bydel"))
```

## Update KnrHarm

```{r}
# add orig = KnrHarm

KnrHarm[, orig := "KnrHarm"]

# combine kommune and fylke, remove oldCode == NA
# change names to GEO, GEO_omk, and HARMstd
# Add orig = orgdata
orgdatarecode <- rbindlist(list(kommune, fylke))[!is.na(oldCode)][, orig := "orgdata"]
setnames(orgdatarecode, 
         c("oldCode", "currentCode", "changeOccurred"),
         c("GEO", "GEO_omk", "HARMstd"))

# Join the existing KnrHarm and orgdata tables
comb <- rbindlist(list(KnrHarm, orgdatarecode))

# Remove rows where GEO == GEO_omk (which doesn't make sense)
comb <- comb[GEO != GEO_omk]

# Remove rows where GEO_omk reappears in GEO due to future recoding
# e.g. 02 (Akershus) -> 30 (Viken) -> 32 (Akershus)
validrecode <- comb[!GEO_omk %in% GEO][order(GEO)]

### Check invalid recode, including the current recoding of GEO and GEO_omk
invalidrecode <- comb[GEO_omk %in% GEO]
invalidrecode[validrecode, correct_omk := i.GEO_omk, on = .(GEO = GEO)]
invalidrecode[validrecode, subsequent_omk := i.GEO_omk, on = .(GEO_omk = GEO)]
invalidrecode[, c("HARMstd", "orig") := NULL]

# Only keep unique combinations of GEO and GEO_omk
validrecode <- unique(validrecode, by = c("GEO", "GEO_omk"))

# Check that all values in GEO_omk are valid
# validcodes are fetched from tblGeo, where validTo = year
validcodes <- tblGeo[validTo == year, (code)]
if(nrow(validrecode[!GEO_omk %in% validcodes]) > 1){
    message("the following rows contain invalid values in GEO_omk")
    validrecode[!GEO_omk %in% validcodes]
} else {
    message("All values in GEO_omk are valid for ", year)
}


# TODO!!!!
# Rader som er fjernet fra KnrHarm og ikke tatt med i endelig tabell
# Trenger å opprette rader for GEO -> correct_omk
removed <- KnrHarm[!GEO %in% validrecode$GEO]

invalidrecode[GEO %in% removed$GEO]

kommune[oldCode == 3801]

```

## Update GeoKoder

```{r}

```
